---
title: "How to Use SUMSarizer"
output: 
 rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Use SUMSarizer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)

knitr::opts_knit$set(
  style="max-width: 1000px",
  width=1000
)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(width = 20)
```

SUMSarizer is an R library that can be used to help analyze 

# Installation
```{r installation, eval=FALSE}
library(devtools)
install_github("geocene/sumsarizer")
```

# Setting up R

```{r setup}
library(sumsarizer)
library(tools)
library(data.table)
# tmp_path <- tempdir()
tmp_path <- "~/tmp"
example_data_path <- file.path(tmp_path, "example_data")
```

# Importing SUMS data

```{r download_example, eval=FALSE}
  download_example_data(example_data_path)
```


```{r import_sums}
  example_sums_file <- file.path(example_data_path, "raw_sums_files","alfashir1_B12.csv")
  one_sums <- import_sums(example_sums_file)
```

# Applying Default SUMSarizer event detection functions
## Threshold Detector
```{r do_threshold_detector}
  one_sums_thresholded <- apply_detector(one_sums, threshold_detector)
```

## FireFinder Detector
```{r do_firefinder_detector}
  one_sums_firefinder <- apply_detector(one_sums, firefinder_detector)
```


## Summarize Detected Events

```{r summarize_events}
  events <- list_events(one_sums_firefinder)
```


We can see the first few events here:

```{r summarize_events_table, output="asis", echo=FALSE}
  kable(head(events))
```

```{r plot_events, fig.width=10, fig.height=6, fig.align="center"}
  plot_sums(one_sums_firefinder)
```




# Creating a custom ensemble ML model

## Exporting SUMS data to TRAINSET for labeling


```{r do_export_trainset, eval=FALSE}
  raw_sums_path <- file.path(example_data_path, "raw_sums_files")
  trainset_path <- file.path(example_data_path, "trainset_files")
  raw_sums_to_trainset(raw_sums_path, trainset_path)
```

## Importing labeled data from TRAINSET

```{r do_import_trainset}
  labeled_path <- file.path(example_data_path, "labeled_files")
  labeled_data <- import_folder(labeled_path)
```

## Choosing models for the learner

See the [sl3 Introductory materials](https://tlverse.org/tlverse-handbook/ensemble-machine-learning.html)

## Training the model

```{r train_sl3_model}
model_obj <- learn_labels(labeled_data)
```


## Storing the model object

```{r save_sl3_model}
model_file <- file.path(tmp_path,"sumsarizer_model_fit.rdata")
save(model_obj, file=model_file)
```

# Using a trained model


## Using the trained model to predict cooking

```{r do_sl3_detector, fig.width=10, fig.height=6, fig.align="center"}
one_sums_ml <- apply_detector(one_sums, sl3_model_detector, model_obj)
plot_sums(one_sums_ml)
```

## Thresholding predictions

```{r do_sl3_detector_options, fig.width=10, fig.height=6, fig.align="center"}
one_sums_ml_sensitive <- apply_detector(one_sums, sl3_model_detector, model_obj,0.1)
plot_sums(one_sums_ml_sensitive)
```


