---
title: "How to Use SUMSarizer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Use SUMSarizer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

SUMSarizer is an R library that can be used to help analyze 

# Installation
```{r installation, eval=FALSE}
library(devtools)
install_github("geocene/sumsarizer")
```

# Setting up R

```{r setup}

library(sumsarizer)
library(tools)
library(data.table)
```

# Importing SUMS data

```{r download_example, eval=FALSE}
  download_example_data("~/tmp/example_data")
```


```{r import_sums}
#TODO
```

# Applying Default SUMSarizer event detection functions
## Threshold Detector
```{r do_threshold_detector}
#TODO
```

## FireFinder Detector
```{r do_firefinder_detector}
#TODO
```


## Summarize Detected Events



# Creating a custom ensemble ML model

## Exporting SUMS data to TRAINSET for labeling


```{r do_export_trainset}
  raw_sums_path <- "~/tmp/example_data/raw_sums_files/"
  trainset_path <- "~/tmp/example_data/trainset_files/"
  raw_sums_to_trainset(raw_sums_path, trainset_path)
```

## Importing labeled data from TRAINSET

```{r do_import_trainset}
#TODO
```

## Choosing models for the learner

```{r define_sl3_model}
#TODO
```

## Training the model

```{r train_sl3_model}
#TODO
```


## Storing the model object

```{r train_sl3_model}
#TODO
```

# Using a trained model


## Using the trained model to predict cooking

```{r do_sl3_detector}
#TODO
```

## Thresholding predictions

```{r do_sl3_detector_options}
#TODO
```


