---
title: "How to Use SUMSarizer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Use SUMSarizer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

SUMSarizer is an R library that can be used to help analyze 

# Setting up R

```{r setup}
library(devtools)
install_github("geocene/sumsarizer")
library(sumsarizer)
library(tools)
library(data.table)
```

# Importing SUMS data

# Applying Default SUMSarizer event detection functions
## Threshold Detector
## Constant Detector
## FireFinder Detector
## Pre-Trained ML Detector

# Creating a custom ensemble ML model
## Exporting SUMS data to TRAINSET for labeling
## Importing labeled data from TRAINSET
## Choosing models for the learner
## Training the model
## Storing the model object

# Using a trained model
## Using the trained model to predict cooking
## Thresholding predictions
## Creating events from thresholded predictions


